

##################################################
### Gardiner Spillway Q vs Turbine Q 2011 

# stringr can be used when need to calculate and concatenate new variables simultaneously

# read file
# cntrl + alt + i to insert chinks of code

```{r}

infile <- 'GardinerQ_SpillwayPower.csv'
data1 <- read.table(
  file = infile,
  header = TRUE,
  sep = ',',
  stringsAsFactors = FALSE
)

names(data1)


library(dplyr)
library(stringr)
library(ggplot2)


```
# plot spillway vs turbine outflows for 3 calibration years (2011-2013)
# try plotting solid vs dotted lines of same colour for each year

```{r}


### maximum

# make decade a factor so can be used as colour variable
data1$Year <- as.factor(data1$Year)


# check dates are formatted properly
data1$Dates <- as.Date(data1$Dates, format="%d/%m/%Y")


# data1$Dates <- as.numeric(data1$Dates, format="%d/%m/%Y")

# check levels of factor
# levels alone only works on vectors so specify df and which column
levels(data1[,4])

## have to manipulate dataframe so have type of withdrawal as a factor and Q as 
## corresponding variable - try dplyr

  # data2 <-data1 %>%
  # group_by(Total_Q,Power_House,Spillway)

# try melt
library(reshape2)

data2 <- data.frame(melt (data1, id=c("Dates","Month","Day","Year","Total_Q")))
names(data2)
data2$Dates

print.data.frame(data2)

# name new variable (outflow) and value (Q)
data2$Outflow<- as.factor(data2$variable)
levels(data2$Outflow)

data2$Q <- as.numeric(data2$value)
  
# make day variable to plot by day of the year

data2$dayoy <-as.numeric(format(data2$Dates, "%j"))
names(data2)
data2$dayoy

# Plot Q by day of year, colour by year and lines by outflow type
# still busy, add plots for turnover periods and stratification

tiff(filename="Spillway_Powerhouse_Q_years.tiff", width=1000, height=600)

ggplot()+
  geom_line(data=data2, mapping = aes(x=dayoy, y=Q, colour= Year, linetype=Outflow))+
  scale_x_continuous(breaks=seq(0,370,20))+
  labs(y="Discharge (cms)")+
labs(x="Day of Year")+
theme(plot.title= element_text(hjust=0.5),panel.grid.major=element_blank(), 
panel.grid.minor=element_blank(),panel.background=element_blank(),
panel.border = element_rect(fill=NA,colour = "black"),
axis.text.x=element_text(colour="black"),axis.text.y=element_text(colour="black")
)



```
######## Spillway vs Turbine outflow Spring Turnover ########
############################################################

```{r}
#### only plot April 15- May 31st (biggest interval = days 105-152)

# make a new month/day variablew to use as x (make based on year 2000)

data2$CommonDate <- as.Date(paste0("2000-", format(data2$Date, "%j")), "%Y-%j")

# subset data2 dataframe

sto <-data2[data2$dayoy > 104 & data2$dayoy < 153,]

tiff(filename="Spillway_Powerhouse_Q_SpringTO.tiff", width=1000, height=600)

ggplot()+
  geom_line(data=sto, mapping = aes(x=CommonDate, y=Q, colour= Year, linetype=Outflow),size=1.2)+
  labs(y="Discharge (cms)")+
labs(x="Day of Year")+
   scale_x_date(labels= function(x) format (x, "%d-%b"))+
theme(plot.title= element_text(hjust=0.5),panel.grid.major=element_blank(), 
panel.grid.minor=element_blank(),panel.background=element_blank(),
panel.border = element_rect(fill=NA,colour = "black"),
axis.text.x=element_text(colour="black"),axis.text.y=element_text(colour="black")
)


```

######### Spillway vs Turbine outflow Fall Turnover ##########
##############################################################

```{r}
#### only plot Sept 15- Oct 31st (biggest interval = days 258-305)

# make a new month/day variablew to use as x (make based on year 2000)

data2$CommonDate <- as.Date(paste0("2000-", format(data2$Date, "%j")), "%Y-%j")

# subset data2 dataframe

sto <-data2[data2$dayoy > 258 & data2$dayoy < 305,]

tiff(filename="Spillway_Powerhouse_Q_FallTO.tiff", width=1000, height=600)

ggplot()+
  geom_line(data=sto, mapping = aes(x=CommonDate, y=Q, colour= Year, linetype=Outflow),size=1.2)+
  labs(y="Discharge (cms)")+
labs(x="Day of Year")+
   scale_x_date(labels= function(x) format (x, "%d-%b"))+
theme(plot.title= element_text(hjust=0.5),panel.grid.major=element_blank(), 
panel.grid.minor=element_blank(),panel.background=element_blank(),
panel.border = element_rect(fill=NA,colour = "black"),
axis.text.x=element_text(colour="black"),axis.text.y=element_text(colour="black")
)

```
######### Spillway vs Turbine outflow Summer Stratification ##########
######################################################################

```{r}

#### only plot June 1- Aug 30 (biggest interval = days 258-305)

# make a new month/day variablew to use as x (make based on year 2000)

data2$CommonDate <- as.Date(paste0("2000-", format(data2$Date, "%j")), "%Y-%j")

# subset data2 dataframe

sto <-data2[data2$dayoy > 151 & data2$dayoy < 305,]

tiff(filename="Spillway_Powerhouse_Q_SummerStrat.tiff", width=1000, height=600)

ggplot()+
  geom_line(data=sto, mapping = aes(x=CommonDate, y=Q, colour= Year, linetype=Outflow),size=1.2)+
  labs(y="Discharge (cms)")+
labs(x="Day of Year")+
   scale_x_date(labels= function(x) format (x, "%d-%b"))+
theme(plot.title= element_text(hjust=0.5),panel.grid.major=element_blank(), 
panel.grid.minor=element_blank(),panel.background=element_blank(),
panel.border = element_rect(fill=NA,colour = "black"),
axis.text.x=element_text(colour="black"),axis.text.y=element_text(colour="black")
)

```

